Herring: case study 1

### 2016-10-14 diversity/analysis/herring_case_study.Rmd
update

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
library(dplyr)
library(knitr)
library(ggplot2)
library(date)
```

@@

>>>>

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
library(dplyr)
library(knitr)
library(ggplot2)
library(date)
library(reshape2)
library(cluster)
```

```{r}
cfec <- feather::read_feather("../../portfolio/data-generated/cfec.feather")
cfec$year <- as.numeric(cfec$year)

this.permit = "S 03T"
# S01E S03T - effective permit diversity

simp.div = function(x) {
  1/sum((x/sum(x))^2)
}

g1 = group_by(cfec, p_holder) %>% 
  mutate(pws = ifelse(length(which(p_fshy==c(this.permit)))>0, 1, 0)) %>% 
  filter(pws>0) %>%
  select(-pws) %>% 
  group_by(p_holder, year, p_fshy) %>% 
  summarize(g = sum(g_earn)) %>%
  group_by(p_holder, year) %>% 
  summarize(div = simp.div(g)) %>% 
  group_by(year) %>% 
  summarize(mean = mean(div), lower=quantile(div,0.025), upper=quantile(div,0.975)) %>% 
  ggplot(aes(year, mean)) + geom_line() + ggtitle(this.permit) + ylab("Mean diversity")

# Identify dominant strategies associated with this permit
g2 = group_by(cfec, p_holder) %>% 
  mutate(pws = ifelse(length(which(p_fshy==c(this.permit)))>0, 1, 0)) %>% 
  filter(pws>0) %>%
  select(-pws) %>% 
  group_by(p_holder, year, p_fshy) %>% 
  summarize(g = sum(g_earn)) %>%
  group_by(p_holder, year) %>% 
  mutate(totRev = sum(g)) %>% 
  group_by(p_holder, year, p_fshy) %>% 
  mutate(g = g / totRev) %>% 
  select(-totRev)

# use reshape 2 to cast this to wide format
aql <- melt(g2, id.vars = c("year", "p_fshy", "p_holder"))
aqw <- dcast(aql, year + p_holder ~ p_fshy)
# replace NAs w/0s
for(i in 1:ncol(aqw)) {
  aqw[which(is.na(aqw[,i])),i] = 0
}

indx = sample(seq(1,nrow(aqw)),size=500,replace=F)
sub = aqw[indx,-c(1,2)]

sub = aqw[,-c(1,2)]
coarse_df = aqw[,-c(1,2)]
coarse_df$S = apply(coarse_df[,which(substr(names(sub),1,1)=="S")],1,sum)
coarse_df$B = apply(coarse_df[,which(substr(names(sub),1,1)=="B")],1,sum)
coarse_df$GHL = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("G","H","L"))],1,sum)
coarse_df$KTD = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("K","T","D"))],1,sum)
coarse_df$M = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("M"))],1,sum)
coarse_df$P = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("P"))],1,sum)
coarse_df = coarse_df[,c("S","B","GHL","KTD","M","P")]

clust = fpc::pamk(coarse_df[sample(seq(1,nrow(aqw)),size=500,replace=F),])# cluster based on subset to pick medoids
medoids = clust$pamobject$medoids
clust = pam(coarse_df, k = clust$nc) # cluster full dataset

names_medoids = ""
for(i in 1:nrow(medoids)) {
  names_medoids[i] = paste0(colnames(medoids)[which(medoids[i,] > 0.05)], collapse=":")
}

aqw$clusters = clust$clustering
totals = group_by(aqw, clusters, year) %>% 
  summarize(n = n())
totals$group = names_medoids[totals$clusters]

ggplot(totals, aes(year, n, group=group, color=group)) + geom_line() + geom_point()

ggplot(totals, aes(year, n)) + geom_area(aes(fill = group), position = "stack", colour = 1)

```{r}

```{r}
cfec <- feather::read_feather("../../portfolio/data-generated/cfec.feather")
cfec$year <- as.numeric(cfec$year)

# grab people who ever held prince william sound herring permit
# calculate unique permits by person-year
pws.G34 = group_by(cfec, p_holder) %>% 
  mutate(pws = ifelse(length(which(p_fshy==c("G 34E")))>0, 1, 0)) %>% 
  filter(pws>0) %>%
  select(-pws) %>% 
  group_by(p_holder, year) %>% 
  summarize(n = length(unique(p_fshy))) %>%
  group_by(year) %>% 
  summarize(p1 = length(which(n>1)) / length(which(n > 0)))

pws.G01 = group_by(cfec, p_holder) %>% 
  mutate(pws = ifelse(length(which(p_fshy==c("G 01E")))>0, 1, 0)) %>% 
  filter(pws>0) %>%
  select(-pws) %>% 
  group_by(p_holder, year) %>% 
  summarize(n = length(unique(p_fshy))) %>%
  group_by(year) %>% 
  summarize(p1 = length(which(n>1)) / length(which(n > 0)))

pws.G34$permit="g34"
pws.G01$permit="g01"
pws = rbind(pws.G34, pws.G01)

ggplot(pws, aes(year, p1, group=permit, color=permit)) + geom_line() + 
  ylab("Proportion of people landing fish under > 1 permit") + xlab("Year")

```{r}
# grab people who ever held prince william sound herring permit
# calculate unique permits by person-year
pws.G34 = group_by(cfec, p_holder) %>% 
  mutate(pws = ifelse(length(which(p_fshy==c("G 34E")))>0, 1, 0)) %>% 
  filter(pws>0) %>%
  select(-pws) %>% 
  group_by(p_holder, year) %>% 
  summarize(n = length(unique(p_fshy))) %>%
  group_by(year) %>% 
  summarize(p1 = length(which(n>1)) / length(which(n > 0)))

pws.G01 = group_by(cfec, p_holder) %>% 
  mutate(pws = ifelse(length(which(p_fshy==c("G 01E")))>0, 1, 0)) %>% 
  filter(pws>0) %>%
  select(-pws) %>% 
  group_by(p_holder, year) %>% 
  summarize(n = length(unique(p_fshy))) %>%
  group_by(year) %>% 
  summarize(p1 = length(which(n>1)) / length(which(n > 0)))

pws.G34$permit="g34"
pws.G01$permit="g01"
pws = rbind(pws.G34, pws.G01)

ggplot(pws, aes(year, p1, group=permit, color=permit)) + geom_line() + 
  ylab("Proportion of people landing fish under > 1 permit") + xlab("Year")

@@

# grab people who ever held prince william sound herring permit
# calculate unique permits by person-year
pws.G34 = group_by(cfec, p_holder) %>% 
  mutate(pws = ifelse(length(which(p_fshy==c("G 34E")))>0, 1, 0)) %>% 
  filter(pws>0) %>%
  select(-pws) %>% 
  group_by(p_holder, year) %>% 
  summarize(n = length(unique(p_fshy))) %>%
  group_by(year) %>% 
  summarize(p1 = length(which(n>1)) / length(which(n > 0)))

pws.G01 = group_by(cfec, p_holder) %>% 
  mutate(pws = ifelse(length(which(p_fshy==c("G 01E")))>0, 1, 0)) %>% 
  filter(pws>0) %>%
  select(-pws) %>% 
  group_by(p_holder, year) %>% 
  summarize(n = length(unique(p_fshy))) %>%
  group_by(year) %>% 
  summarize(p1 = length(which(n>1)) / length(which(n > 0)))

pws.G34$permit="g34"
pws.G01$permit="g01"
pws = rbind(pws.G34, pws.G01)

ggplot(pws, aes(year, p1, group=permit, color=permit)) + geom_line() + 
  ylab("Proportion of people landing fish under > 1 permit") + xlab("Year")
